validateHOT - an R package for holdout task validation and market simulations

### 2024-03-03 paper/paper.Rmd
updated paper - 240303

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [e.g., @rao2014; @Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities (raw logit utilities) in hierarchical Bayes estimation. Despite the importance of validation/holdout tasks, practitioners do not always include them [@yang2018]. This is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

[^1]: In terms of external validity, we refer to the generalizations to different settings [see, @calder1982, p.240].

validateHOT combines both validation and market simulation in one package and has three key advantages, it (1) helps to opt for the best model and (2) runs relevant market simulations that help to find the right product combinations or assortments, and finally, (3) is an open source tool which helps especially researchers to report accompanied scripts for their research papers.

@@

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] focuses on classical conjoint analysis, and thus is limited when applying more common conjoint methods, for example, (A)CBC. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as MaxDiff) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```
validateHOT is introduced with data estimated with Lighthouse Studio using effects-coding for creating the design matrix. It, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT()} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf()} as well as the four rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT()}, and can be run using the raw logit scores.

@@

: Overview of validateHOT's main four components and their corresponding functions \label{tab:table1}

@@

In the following, we provide the workflow for a MaxDiff study and a CBC study with only part-worth coded attributes (the vignette also provides detailed examples for a CBC including linear-coded attributes as well as an ACBC).

@@

After running the hierarchical Bayes estimation [@allenby1995; @lenk1996], the **raw** utility scores have to be exported and read into an *R* data frame. This data frame has to include the actual choice in the validation/holdout task (if only a market scenario is created, the \texttt{\color{purple}choice} argument in \texttt{\color{purple}createHOT()} can be left empty)..

Assuming you included a validation/holdout task with a total of 7 alternatives plus the no-buy alternative (\texttt{\color{purple}none}). To create this validation task in *R*, we use the \texttt{\color{purple}createHOT()} function.

@@

To get the relevant validation metrics that are reported in conjoint studies, for example, hit rate [e.g., @ding2005], mean hit probability [mhp, @voleti2017], or mean absolute error [mae, @wloemert2014], we provide the data, the alternatives in the validation/holdout task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}), which can be implemented using the tidyverse [@tidyverse] logic.

\newpage

@@

The underlying logic of the confusion matrix is that the user has to provide a no-buy alternative (\texttt{\color{purple}none}). validateHOT calculates how often a buy or no-buy was correctly predicted, therefore, it is testing whether the model correctly predicts general demand (here by applying \texttt{\color{purple}accuracy()}).

```{r}
accuracy(
  data = HOT, 
  group = Group, 
  opts = c(Option_1:None), 
  choice = choice, 
  none = None 
) %>% 
  round(2)
```

@@

Lastly, we introduce two functions for market simulations, namely \texttt{\color{purple}marksim()} and \texttt{\color{purple}turf()}. In the following example, the market share is calculated according to the multinomial logit model [@McFadden1974].

@@

Next, \texttt{\color{purple}turf()}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibit certain item combinations of being part of the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf()} with data obtained using a likert scale).

@@

The setup is almost the same, only the arguments \texttt{\color{purple}prod.levels}, \texttt{\color{purple}coding}, and \texttt{\color{purple}method} are different or new, respectively. \newpage

@@

This time we calculate the mean hit probability (i.e., MHP).

```{r}
HOT_CBC %>% 
  mhp(
  data = ., 
  opts = c(Option_1:None), 
  choice = choice 
) %>% 
  round(2)
```

>>>>

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well. The easiest way for testing validity is by including validation tasks [e.g., @rao2014; @Orme2015], which are fixed tasks (i.e., same across participants) and not used for estimating the part-worth utilities (raw logit utilities) in hierarchical Bayes (HB) estimation. Despite their importance, practitioners don't always include them [@yang2018]. This is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

validateHOT combines both validation and market simulation in one package and has three key advantages, it (1) helps opting for the best model and (2) runs relevant market simulations that help finding the right product combinations or assortments, and (3) is an open source tool which helps especially researchers reporting accompanied scripts for their research papers.

@@

Other packages provide functions to calculate validation metrics, however, these are not always specified for individual part-worth utilities. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones, but no validation functions are included and the package focuses on classical conjoint analysis, thus it is limited when applying more common conjoint methods. support.BWS [@support.BWS] only covers best-worst scaling case 1 (i.e., MaxDiff). logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. \autoref{comparison} shows a comparison of validateHOT's functions with current R packages. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```

validateHOT is introduced with data estimated with Lighthouse Studio using effects-coding for creating the design matrix. It, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format for some functions, the \texttt{\color{purple}createHOT()} function can be applied, which creates each alternatives' total utility by applying the additive utility model [@rao2014, p. 82].

@@

: Overview of validateHOT's main components and their corresponding functions \label{tab:table1}

@@

We provide the workflow for a MaxDiff study and a CBC study with only part-worth coded attributes (the vignette provides detailed examples for other CBCs and an ACBC).

@@

After running the HB estimation [@allenby1995; @lenk1996], the **raw** utility scores have to be exported and read into an *R* data frame. 
Assuming you included a validation task with seven alternatives plus the no-buy alternative (\texttt{\color{purple}none}). To create this validation task in *R*, we use the \texttt{\color{purple}createHOT()} function.

@@

To get the relevant validation metrics that are reported in conjoint studies, for example, hit rate [e.g., @ding2005] or mean hit probability [mhp, @voleti2017], we provide the data, the alternatives in the validation task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). The function can be implemented using the tidyverse [@tidyverse] logic.

@@

@@

We also introduce two functions for market simulations, namely \texttt{\color{purple}marksim()} and \texttt{\color{purple}turf()}. In the following example, the market share is calculated according to the multinomial logit model [@McFadden1974].

@@

Next, \texttt{\color{purple}turf()}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal assortment, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibit certain item combinations in the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf()} with data obtained using a likert scale).

@@

The setup is almost the same, only the arguments \texttt{\color{purple}prod.levels}, \texttt{\color{purple}coding}, and \texttt{\color{purple}method} are different or new, respectively.

@@
